Fast Nearest Neighbour Classification

### 2016-05-11 paper/paper.tex
finished

\title{Fast Nearest Neighbour Classification}
\subtitle{\textnormal{\small{Seminar\\
Intelligente Software-Systeme\\
Sommersemester 2015\\\vspace{1\baselineskip}
Stefan Fricke\\\vspace{2\baselineskip}
April 2016\\
Betreuer: Stephan Spiegel\\\vspace{1\baselineskip}}}}

@@

HERE COMES THE INTRODUCTION.

@@

Every nearest neighbour search algorithm can be divided into two steps.
\paragraph{Preprocessing}
During the preprocessing the algorithm prepares the given set $\mathbb{S}$ for his needs.

@@

The advantage compared to the following algorithms is, that the Full Search works also in not metric spaces. The Full
Search needs no preprocessing.

@@

When calling the query processing with query $q$, the algorithm will randomly select one object $c \in S$ as initial
candidate and calculate $D(c, q)$. Afterwards the algorithm walks over the ordered list of $c$ and with every picked $s$
from the list the algorithm calculates $D(l, q)$. If $D(l, q)$ is smaller than $D(c, q)$, $l$ is the new candidate $c$
and the algorithm walks over the list of $l$. The algorithm aborts if it reaches the end of a list or if
$D(c, s) > 2D(c, q)$. The current candidate $c$ is the nearest neighbour.

@@

There exists a improved version of the Orchard’s Algorithm to ensure that no distance between an item $x \in S$ and $q$
is calculated twice. This can achieved by boolean flags for every item in the query processing of the Orchard’s
Algorithm.

@@

The Annulus Method works a little bit similar to the Orchard’s Algorithm, but with a less complex preprocessing.

@@

Instead of generating a list for all $x \in S$, the Annulus Method generates just one list for a random $p^* \in S$. In
the preprocessing the algorithm generates one list that contains all $x \in S$ with their distance $D(p^*, x)$. This
list will be ordered ascending to the distance. For a set $S$ with size $n$, the preprocessing will call the distance
function $n$ times and will sort one list of size $n$.

@@

The query processing of the Annulus Method starts by picking a random candidate $c \in S$ from the ordered list of
$p^*$. Afterwards the algorithm walks alternating away from $p^*$ and back to him in the list. If the current object $s$
has $D(s, q) < D(c, q)$, the algorithm sets $s$ as the new candidate $c$. If the current object $s$ is under $c$ in the
list and $D(p^*, s) < D(p^*, q) - D(c, q)$, the algorithm can ignore all objects under $s$ in the list. If the current
object $s$ is above $c$ in the list and $D(p^*, s) > D(p^*, q) + D(c, q)$, the algorithm can ignore all objects above
$s$ in the list. The object $c$ is the nearest neighbour, if the entire list is traversed.

@@

When starting with the query processing, the algorithm generates a list for all objects $x \in S$ with a bounded
distance $D_P(x, q) = -\infty$ and initialize $r = \infty$ as smallest distance to $q$. Afterwards the algorithm
iterates the following steps until the list is empty.
\begin{itemize}
	\item pick and remove $x$ from the list with the smallest $D_p(x, q)$ or a random one in the first iteration
	\item calculate $D(x, q)$
	\item set $r = D(x, q)$ if $r > D(x, q)$ and remember $x$ as current nearest neighbour
	\item for all $y$ from the list, set $D_P(y, q) = max(D_P(y, q), |D(x, q) - D(x, y)|)$ and remove $y$ from the list
		if the new $D_P(y, q) > r$
\end{itemize}

@@

As described above, the following benchmarks will not compare the runtime of the mentioned algorithms. The benchmarks
will compare the calls of the distance function that the algorithm needs for preprocessing and for finding a nearest
neighbour. Those benchmarks can make sense when the distance function has a heavy complexity. For example the string
metric Levenshtein distance or in higher dimensional vector spaces.\\
In general a benchmark is affected by the implementation, by the programming language, by the executing machine, by the
domain of the problem and the distribution of the example. As the following benchmarks only counting the calls of the
distance function, they are not affected by the programming language or the executing machine.

@@

The algorithms above are all implemented in Java and can be found in the following git
repository\footnote[1]{\url{https://github.com/GordonLesti/FastNearestNeighbourClassification}}.\\
Since the benchmakrs are independed from the programming language and the executing machine, they use sometimes not the
most effective data structures, but that never effects the logical flow of the algorithms.\\

A nearest neighbour algorithm implemented in Java should handle every type of object that a set $\mathbb{U}$ can
represent under the condition that there is a metric on $\mathbb{U}$. Because of that, a nearest neighbour algorithm
class needs to extend from the abstract class\\ \verb+FastNearestNeighbourClassificator<T, D extends Comparable<D>>+
that\\ can be found in the package \verb+fnnc.algo+. The generic \verb+T+ represents an object from the set $\mathbb{U}$
and the generic \verb+D+ represents the distance and has to be of type \verb+Comparable+ from the package
\verb+java.lang+.\\

@@

A instance of type \verb+FastNearestNeighbourClassificator+ should not be allowed and can not calculate the distance
between two objects. So that a instance of \verb+DistanceCalculator+ should be injected into the constructor. All
implemented algorithms exclude the \verb+FullSearch+ require that the\\ \verb+DistanceCalculator+ returns a distance of
type \verb+Double+ instead of only\\ \verb+Comparable+. This is connected with the fact that only the Full Search can
operate in a not metric space.

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{| l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	$n$ & 0 & $\frac{n(n-1)}{2}$ & $\frac{n(n-1)}{2}$ & $n$ & $\frac{n(n-1)}{2}$ \\ \hline
		\end{tabular}
	\end{center}
	\caption{The calls of the distance function $D$ in the preprocessing depending on the size $n$ of the test data set
		$S$.}
	\label{fig:preprocessing:tabular}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[
				ymin=0,
				xmin=0,
				xmax=2000,
				xlabel=size of test data set $S$,
				ylabel=calls of distance function $D$,
				legend pos=outer north east]
				\addplot table {data/preprocessing/fullsearch.csv};
	            \addlegendentry{Full Search}
	            \addplot table {data/preprocessing/orchards.csv};
	            \addlegendentry{Orchard}
				\addplot table {data/preprocessing/orchardsmb.csv};
				\addlegendentry{Orchard MarkBits}
	            \addplot table {data/preprocessing/annulus.csv};
	            \addlegendentry{Annulus}
	            \addplot table {data/preprocessing/aesa.csv};
	            \addlegendentry{AESA}
			\end{axis}
		\end{tikzpicture}
	\end{center}
	\caption{The average calls of the distance function $D$ in the preprocessing depending on the size of the test data
		set $S$ visualized with graphs.}
	\label{fig:preprocessing:graph}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{| l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	100 & 100 & 29 & 20 & 46 & 4 \\ \hline
			200 & 200 & 39 &26 & 82 & 4 \\ \hline
			300 & 300 & 46 & 30 & 117 & 4 \\ \hline
			400 & 400 & 53 & 35 & 160 & 4 \\ \hline
			500 & 500 & 57 & 40 & 189 & 4 \\ \hline
			600 & 600 & 63 & 42 & 228 & 4 \\ \hline
			700 & 700 & 68 & 44 & 263 & 4 \\ \hline
			800 & 800 & 73 & 46 & 303 & 4 \\ \hline
			900 & 900 & 78 & 50 & 341 & 4 \\ \hline
			1000 & 1000 & 79 & 51 & 367 & 4 \\ \hline
			1100 & 1100 & 84 & 56 & 409 & 4 \\ \hline
			1200 & 1200 & 86 & 56 & 442 & 4 \\ \hline
			1300 & 1300 & 89 & 59 & 466 & 4 \\ \hline
			1400 & 1400 & 90 & 60 & 510 & 4 \\ \hline
			1500 & 1500 & 97 & 63 & 541 & 4 \\ \hline
			1600 & 1600 & 100 & 66 & 573 & 4 \\ \hline
			1700 & 1700 & 102 & 67 & 624 & 4 \\ \hline
			1800 & 1800 & 104 & 69 & 660 & 4 \\ \hline
			1900 & 1900 & 109 & 71 & 664 & 4 \\ \hline
			2000 & 2000 & 113 & 73 & 740 & 4 \\ \hline
		\end{tabular}
	\end{center}
	\caption{The average calls of the distance function $D$ in the query processing depending on the size of the test
		data set $S$.}
	\label{fig:queryprocessing:tabular}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[
				ymin=0,
				xmin=0,
				xmax=2000,
				xlabel=size of test data set $S$,
				ylabel=average calls of distance function $D$,
				legend pos=outer north east]
				\addplot table {data/queryprocessing/fullsearch.csv};
	            \addlegendentry{Full Search}
	            \addplot table {data/queryprocessing/orchards.csv};
	            \addlegendentry{Orchard}
				\addplot table {data/queryprocessing/orchardsmb.csv};
				\addlegendentry{Orchard MarkBits}
	            \addplot table {data/queryprocessing/annulus.csv};
	            \addlegendentry{Annulus}
	            \addplot table {data/queryprocessing/aesa.csv};
	            \addlegendentry{AESA}
			\end{axis}
		\end{tikzpicture}
	\end{center}
	\caption{The average calls of the distance function $D$ in the query processing depending on the size of the test
		data set $S$ visualized with graphs.}
	\label{fig:queryprocessing:graph}
\end{figure}

@@

As I saw that some of the algorithms needed more and more tries to stabilize the average result, I have calculated for
every size of a set and every algorithm the standard deviation. In general it is no big surprise that algorithms that
start with random initialization have fluctuating results, but especially the Annulus Method was very fragile.

\begin{figure}
	\begin{center}
		\begin{tabular}{| l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	100 & 0 & 13 & 8 & 23 & 1 \\ \hline
			200 & 0 & 17 & 11 & 46 & 1 \\ \hline
			300 & 0 & 21 & 13 & 69 & 1 \\ \hline
			400 & 0 & 25 & 15 & 94 & 1 \\ \hline
			500 & 0 & 27 & 17 & 119 & 1 \\ \hline
			600 & 0 & 29 & 19 & 144 & 1 \\ \hline
			700 & 0 & 32 & 19 & 162 & 1 \\ \hline
			800 & 0 & 33 & 21 & 190 & 1 \\ \hline
			900 & 0 & 36 & 22 & 213 & 1 \\ \hline
			1000 & 0 & 37 & 23 & 232 & 1 \\ \hline
			1100 & 0 & 38 & 24 & 264 & 1 \\ \hline
			1200 & 0 & 40 & 26 & 277 & 1 \\ \hline
			1300 & 0 & 43 & 27 & 301 & 1 \\ \hline
			1400 & 0 & 43 & 27 & 339 & 1 \\ \hline
			1500 & 0 & 47 & 27 & 350 & 1 \\ \hline
			1600 & 0 & 46 & 29 & 374 & 1 \\ \hline
			1700 & 0 & 48 & 29 & 401 & 1 \\ \hline
			1800 & 0 & 49 & 31 & 429 & 1 \\ \hline
			1900 & 0 & 52 & 32 & 442 & 1 \\ \hline
			2000 & 0 & 52 & 33 & 487 & 1 \\ \hline
		\end{tabular}
	\end{center}
	\caption{The standard deviation of the calls of the distance function $D$ in the query processing depending on the
		size of the test data set $S$.}
	\label{fig:standarddeviation:tabular}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[
				ymin=0,
				xmin=0,
				xmax=2000,
				xlabel=size of test data set $S$,
				ylabel=standard deviation of calls of distance function $D$,
				legend pos=outer north east]
				\addplot table {data/standarddeviation/fullsearch.csv};
	            \addlegendentry{Full Search}
	            \addplot table {data/standarddeviation/orchards.csv};
	            \addlegendentry{Orchard}
				\addplot table {data/standarddeviation/orchardsmb.csv};
				\addlegendentry{Orchard MarkBits}
	            \addplot table {data/standarddeviation/annulus.csv};
	            \addlegendentry{Annulus}
	            \addplot table {data/standarddeviation/aesa.csv};
	            \addlegendentry{AESA}
			\end{axis}
		\end{tikzpicture}
	\end{center}
	\caption{The standard deviation of the calls of the distance function $D$ in the query processing depending on the
		size of the test data set $S$ visualized with graphs.}
	\label{fig:standarddeviation:graph}
\end{figure}

@@

\paragraph{Full Search} The results may leave the impression that the Full Search is not a good choice. But this
algorithm remains useful if the set $S$ changes completely or when $\mathbb{U}$ is no metric space.

\paragraph{Annulus Method} The Annulus Method has a very complex query processing that

\paragraph{Orchard’s Algorithm with and without marking bits} In the query processing the Orchard’s Algorithm with
marking bits consumed on average only 66\% of the distance function calls that the algorithm needed without marking
bits.

@@

>>>>

\title{Fast Nearest Neighbour Classification}
\subtitle{\textnormal{\small{Seminar\\
Intelligente Software-Systeme\\
Sommersemester 2015\\\vspace{1\baselineskip}
Stefan Fricke\\\vspace{2\baselineskip}
May 2016\\
Supervisor: Stephan Spiegel\\\vspace{1\baselineskip}}}}

@@

Kenneth L. Clarkson wrote the paper \textit{Nearest-Neighbor Searching and Metric Space Dimensions}[1] which deals
amongst other approaches with the topic nearest neighbour search algorithms that use the triangle inequality. This paper
focuses on three of them.

@@

Every nearest neighbour search algorithm can be divided into two steps.
\paragraph{Preprocessing}
During the preprocessing the algorithm prepares the given set $\mathbb{S}$ for its needs.

@@

The advantage compared to the following algorithms is, that the Full Search works also in not metric spaces. The Full
Search does not need a preprocessing.

@@

When calling the query processing with query $q$, the algorithm will randomly select one item $c \in S$ as initial
candidate and calculate $D(c, q)$. Afterwards the algorithm walks over the ordered list of $c$ and with every picked $s$
from the list the algorithm calculates $D(l, q)$. Item $l$ is the new candidate $c$ if $D(l, q)$ is smaller than
$D(c, q)$. Thereafter the algorithm walks over the list of $l$. The algorithm interupts if it reaches the end of a list
or if $D(c, s) > 2D(c, q)$. The current candidate $c$ is the nearest neighbour.

@@

There is a improved version of the Orchard’s Algorithm to ensure that no distance between an item $x \in S$ and $q$
is calculated twice. This can be achieved by boolean flags for every item during query processing of the Orchard’s
Algorithm.

@@

@@

Instead of generating a list for all $x \in S$, the Annulus Method generates just one list for a random $p^* \in S$.
During preprocessing the algorithm generates one list which contains all $x \in S$ with their distance $D(p^*, x)$. This
list will be ordered ascending to the distance. For a set $S$ with size $n$, the preprocessing will call the distance
function $n$ times and will sort one list of size $n$.

@@

The query processing of the Annulus Method starts by picking a random candidate $c \in S$ from the ordered list of
$p^*$. Afterwards the algorithm walks alternating away from $p^*$ and back to it in the list. If the current item $s$
has $D(s, q) < D(c, q)$, the algorithm sets $s$ as the new candidate $c$. If the current item $s$ is under $c$ in the
list and $D(p^*, s) < D(p^*, q) - D(c, q)$, all items under $s$ in the list will be ignored. If the current
item $s$ is above $c$ in the list and $D(p^*, s) > D(p^*, q) + D(c, q)$, the algorithm will ignore all items above
$s$ in the list. The item $c$ is the nearest neighbour, if the entire list is traversed.

@@

When starting with the query processing, the algorithm generates a list for all items $x \in S$ with a bounded
distance $D_P(x, q) = -\infty$ and initialize $r = \infty$ as smallest distance to $q$. Afterwards the algorithm
iterates the following steps until the list is empty.
\begin{itemize}
	\item pick and remove $x$ from the list with the smallest $D_p(x, q)$ or a random one during first iteration
	\item calculate $D(x, q)$
	\item set $r = D(x, q)$ if $r > D(x, q)$ and remember $x$ as current nearest neighbour
	\item for all $y$ from the list, set $D_P(y, q) = max(D_P(y, q), |D(x, q) - D(x, y)|)$ and remove $y$ from the list
		if the new $D_P(y, q) > r$
\end{itemize}

@@

As described above, the following benchmarks will not compare the runtime of the mentioned algorithms. The benchmarks
will compare the calls of the distance function that the algorithm needs for preprocessing and for finding a nearest
neighbour. Those benchmarks are useful when the distance function has a heavy complexity. For example the string
metric Levenshtein distance or in higher dimensional vector spaces.\\
In general a benchmark is affected by the implementation, by the programming language, by the executing machine, by the
domain of the problem and the distribution of the example. As the following benchmarks only counting the calls of the
distance function its are not affected by the programming language or the executing machine.

@@

The algorithms above are all implemented in Java and can be found in the following git
repository\footnote[1]{\url{https://github.com/GordonLesti/FastNearestNeighbourClassification}}.\\
Since the benchmarks are independed from the programming language and the executing machine, the implementation of the
algorithms uses sometimes not the most effective data structures. But that never effects the logical flow of the
algorithms.\\

A nearest neighbour algorithm implemented in Java should handle every type of objects that a set $\mathbb{U}$ can
represent under the condition that there is a metric on $\mathbb{U}$. Because of that, a nearest neighbour algorithm
class needs to extend from the abstract class\\ \verb+FastNearestNeighbourClassificator<T, D extends Comparable<D>>+
that\\ can be found in the package \verb+fnnc.algo+. The generic \verb+T+ represents an object from the set $\mathbb{U}$
and the generic \verb+D+ represents the distance and has to be of type \verb+Comparable+ from the package
\verb+java.lang+.\\

@@

An instance of type \verb+FastNearestNeighbourClassificator+ should not be allowed and can not calculate the distance
between two objects. So that an instance of \verb+DistanceCalculator+ should be injected into the constructor. All
implemented algorithms exclude the \verb+FullSearch+ require that the\\ \verb+DistanceCalculator+ returns a distance of
type \verb+Double+ instead of only\\ \verb+Comparable+. This is connected with the fact that only the Full Search can
operate in a not metric space.

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{| l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	$n$ & 0 & $\frac{n(n-1)}{2}$ & $\frac{n(n-1)}{2}$ & $n$ & $\frac{n(n-1)}{2}$ \\ \hline
		\end{tabular}
	\end{center}
	\caption{The calls of the distance function $D$ during preprocessing depending on the size $n$ of the test data set
		$S$.}
	\label{fig:preprocessing:tabular}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[
				ymin=0,
				xmin=0,
				xmax=2000,
				xlabel=size of test data set $S$,
				ylabel=calls of distance function $D$,
				legend pos=outer north east]
				\addplot table {data/preprocessing/fullsearch.csv};
	            \addlegendentry{Full Search}
	            \addplot table {data/preprocessing/orchards.csv};
	            \addlegendentry{Orchard}
				\addplot table {data/preprocessing/orchardsmb.csv};
				\addlegendentry{Orchard MarkBits}
	            \addplot table {data/preprocessing/annulus.csv};
	            \addlegendentry{Annulus}
	            \addplot table {data/preprocessing/aesa.csv};
	            \addlegendentry{AESA}
			\end{axis}
		\end{tikzpicture}
	\end{center}
	\caption{The average calls of the distance function $D$ during preprocessing depending on the size of the test data
		set $S$ visualized with graphs.}
	\label{fig:preprocessing:graph}
\end{figure}

@@

Figure~\ref{fig:queryprocessing:tabular} and figure~\ref{fig:queryprocessing:graph} show the results of the query
processing for all algorithms.

\begin{figure}
	\begin{center}
		\begin{tabular}{| l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	100 & 100 & 29 & 20 & 46 & 4 \\ \hline
			200 & 200 & 39 &26 & 82 & 4 \\ \hline
			300 & 300 & 46 & 30 & 117 & 4 \\ \hline
			400 & 400 & 53 & 35 & 160 & 4 \\ \hline
			500 & 500 & 57 & 40 & 189 & 4 \\ \hline
			600 & 600 & 63 & 42 & 228 & 4 \\ \hline
			700 & 700 & 68 & 44 & 263 & 4 \\ \hline
			800 & 800 & 73 & 46 & 303 & 4 \\ \hline
			900 & 900 & 78 & 50 & 341 & 4 \\ \hline
			1000 & 1000 & 79 & 51 & 367 & 4 \\ \hline
			1100 & 1100 & 84 & 56 & 409 & 4 \\ \hline
			1200 & 1200 & 86 & 56 & 442 & 4 \\ \hline
			1300 & 1300 & 89 & 59 & 466 & 4 \\ \hline
			1400 & 1400 & 90 & 60 & 510 & 4 \\ \hline
			1500 & 1500 & 97 & 63 & 541 & 4 \\ \hline
			1600 & 1600 & 100 & 66 & 573 & 4 \\ \hline
			1700 & 1700 & 102 & 67 & 624 & 4 \\ \hline
			1800 & 1800 & 104 & 69 & 660 & 4 \\ \hline
			1900 & 1900 & 109 & 71 & 664 & 4 \\ \hline
			2000 & 2000 & 113 & 73 & 740 & 4 \\ \hline
		\end{tabular}
	\end{center}
	\caption{The average calls of the distance function $D$ during query processing depending on the size of the test
		data set $S$.}
	\label{fig:queryprocessing:tabular}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[
				ymin=0,
				xmin=0,
				xmax=2000,
				xlabel=size of test data set $S$,
				ylabel=average calls of distance function $D$,
				legend pos=outer north east]
				\addplot table {data/queryprocessing/fullsearch.csv};
	            \addlegendentry{Full Search}
	            \addplot table {data/queryprocessing/orchards.csv};
	            \addlegendentry{Orchard}
				\addplot table {data/queryprocessing/orchardsmb.csv};
				\addlegendentry{Orchard MarkBits}
	            \addplot table {data/queryprocessing/annulus.csv};
	            \addlegendentry{Annulus}
	            \addplot table {data/queryprocessing/aesa.csv};
	            \addlegendentry{AESA}
			\end{axis}
		\end{tikzpicture}
	\end{center}
	\caption{The average calls of the distance function $D$ during query processing depending on the size of the test
		data set $S$ visualized with graphs.}
	\label{fig:queryprocessing:graph}
\end{figure}

@@

Some of the algorithms needed more and more tries to stabilize the average result of the query processing.
Figure~\ref{fig:standarddeviation:tabular} and figure~\ref{fig:standarddeviation:graph} show the standard deviation of
the results for every algorithm depending on the size of set $S$.

\begin{figure}
	\begin{center}
		\begin{tabular}{| l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	100 & 0 & 13 & 8 & 23 & 1 \\ \hline
			200 & 0 & 17 & 11 & 46 & 1 \\ \hline
			300 & 0 & 21 & 13 & 69 & 1 \\ \hline
			400 & 0 & 25 & 15 & 94 & 1 \\ \hline
			500 & 0 & 27 & 17 & 119 & 1 \\ \hline
			600 & 0 & 29 & 19 & 144 & 1 \\ \hline
			700 & 0 & 32 & 19 & 162 & 1 \\ \hline
			800 & 0 & 33 & 21 & 190 & 1 \\ \hline
			900 & 0 & 36 & 22 & 213 & 1 \\ \hline
			1000 & 0 & 37 & 23 & 232 & 1 \\ \hline
			1100 & 0 & 38 & 24 & 264 & 1 \\ \hline
			1200 & 0 & 40 & 26 & 277 & 1 \\ \hline
			1300 & 0 & 43 & 27 & 301 & 1 \\ \hline
			1400 & 0 & 43 & 27 & 339 & 1 \\ \hline
			1500 & 0 & 47 & 27 & 350 & 1 \\ \hline
			1600 & 0 & 46 & 29 & 374 & 1 \\ \hline
			1700 & 0 & 48 & 29 & 401 & 1 \\ \hline
			1800 & 0 & 49 & 31 & 429 & 1 \\ \hline
			1900 & 0 & 52 & 32 & 442 & 1 \\ \hline
			2000 & 0 & 52 & 33 & 487 & 1 \\ \hline
		\end{tabular}
	\end{center}
	\caption{The standard deviation of the calls of the distance function $D$ during query processing depending on the
		size of the test data set $S$.}
	\label{fig:standarddeviation:tabular}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[
				ymin=0,
				xmin=0,
				xmax=2000,
				xlabel=size of test data set $S$,
				ylabel=standard deviation of calls of distance function $D$,
				legend pos=outer north east]
				\addplot table {data/standarddeviation/fullsearch.csv};
	            \addlegendentry{Full Search}
	            \addplot table {data/standarddeviation/orchards.csv};
	            \addlegendentry{Orchard}
				\addplot table {data/standarddeviation/orchardsmb.csv};
				\addlegendentry{Orchard MarkBits}
	            \addplot table {data/standarddeviation/annulus.csv};
	            \addlegendentry{Annulus}
	            \addplot table {data/standarddeviation/aesa.csv};
	            \addlegendentry{AESA}
			\end{axis}
		\end{tikzpicture}
	\end{center}
	\caption{The standard deviation of the calls of the distance function $D$ during query processing depending on the
		size of the test data set $S$ visualized with graphs.}
	\label{fig:standarddeviation:graph}
\end{figure}

@@

\paragraph{Full Search} The results may leave the impression that the Full Search is not a good choice. But this
algorithm remains useful if the set $S$ changes completely with every new query or when $\mathbb{U}$ is no metric space.

\paragraph{Annulus Method} The Annulus Method has a very complex query processing. But on a frequently changing set $S$
a affordable preprocessing. Other algorithms with more complex preprocessing maybe hit its limits on huge sets that
can be handled by the Annulus Method. The Annulus Method was the algorithm with the highest standard deviation during
the benchmarks.

\paragraph{Orchard’s Algorithm with and without marking bits} During query processing the Orchard’s Algorithm with
marking bits consumed on average only 66\% of the distance function calls that the algorithm needed without marking
bits.

\paragraph{AESA} Depending on the benchmarks above, the AESA should be preferred instead of the Orchard’s Algorithm.
Both algorithms have a very complex preprocessing that may not work on very huge sets. But the AESA consumes more less
calls of the distance function. The impact on real world applications should be tested. Beside the Full Search the AESA
was the most stable algorithm.

@@

\begin{thebibliography}{4}
	\bibitem{url} Kenneth L. Clarkson, \"Nearest-Neighbor Searching and Metric Space Dimensions\" (October 2005), \url{http://kenclarkson.org/nn_survey/p.pdf}
\end{thebibliography}
